# Data Preparation {#sec-data-preparation}

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(scales)
library(janitor)
library(survey)
library(srvyr)
library(gtsummary)
```

This book uses the `api` datasets from the *survey* package for examples. The Academic Performance Index (API) was a school rating system used in California for several years. The *survey* package includes several datasets that mimic survey samples from the overall population of 6,194 schools.

`data(api)` returns several data files: a simple random sample (`apisrs`), a stratified simple random sample (`apistrat`), and a two-stage cluster (`apiclus2`).

```{r}
data(api, package = "survey")

# Add some cols for stat test examples.
prep_data <- function(df) {
  df |>
    mutate(
      stype = factor(stype, levels = c("E", "M", "H"), ordered = TRUE), 
      meals_cut = cut(meals, c(0, 12, 25, 100), include.lowest = TRUE),
      hsg_cut = cut(hsg, c(0, 12, 25, 100), include.lowest = TRUE)
    )
}

apisrs <- prep_data(apisrs)

apistrat <- prep_data(apistrat)

apiclus2 <- prep_data(apiclus2)
```

Schools are uniquely identified by column `snum`. Schools roll up to districts, `dnum`. Two other columns contain metadata related to the sampling design.

- `fpc`: **finite population correction (FPC)**. The FPC adjusts the variance calculation (@sec-sd-srs). The FPC is important when the sample size is >=5% of the population size. `fpc` equals the size of the population that the respondent is drawn from. For an SRS, that's the _entire_ population of 6,194 schools. For a 2 stage cluster design, that's the second stage population.
- `pw`: **sampling weight**. The sampling weight scales the sample up to the population. Think of it as saying, "this respondent represents `pw` respondents from the total population."

Let's create the design objects.

::: panel-tabset

## Simple Random Sample

`apisrs` is a simple random sample of 200 schools from a population of 6,194, so `fpc` = 6194 and `pw` = 6194 / 200 = 30.97 for all rows.

```{r}
apisrs_des <- as_survey_design(apisrs, weights = pw, fpc = fpc)

summary(apisrs_des)
```

## Stratified

`apistrat` is a sample of 200 schools from a population stratified by school type, `stype`: *E* = Elementary (*n* = 100, `fpc` = 4421, `pw` = `r comma(4421/100, .1)`), *M* = Middle (*n* = 50, `fpc` = 1018, `pw` = `r comma(1018/50, .1)`), and *H* = High School (*n* = 50, `fpc` = 755, `pw` = `r comma(755/50, .1)`). `pw` equals the `fpc` / *n*.

```{r}
apistrat_des <- as_survey_design(apistrat, weights = pw, fpc = fpc, strata = stype)

summary(apistrat_des)
```

## Two-Stage Cluster

`apiclus2` is a two-stage cluster sample of 126 schools within districts. The first stage is random sample of 40 of the 757 school districts (`dnum`). The second stage is a random sample of up to 5 schools (`snum`) from each district. For cluster designs, the cluster ids are specified in the design object from largest to smallest level.

```{r}
apiclus2_des <- as_survey_design(
  apiclus2,
  id = c(dnum, snum),
  weights = pw,
  fpc = c(fpc1, fpc2),
)

summary(apiclus2_des)
```

:::

```{r include=FALSE}
save(
  apisrs, apisrs_des, 
  apistrat, apistrat_des,
  apiclus2, apiclus2_des,
  file = "input/api.Rdata"
)
```
