# Modeling {#sec-modeling}

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(scales)
library(janitor)
library(survey)
library(srvyr)
library(gtsummary)
```

```{r include=FALSE}
load("input/api.Rdata")
```

A one-way ANOVA is equivalent to a linear model with a single categorical variable. A two-sample t-test is equivalent to an ANOVA with a two-level categorical predictor.

## ANOVA

ANOVA tests whether the mean outcome is the same across groups. It is equivalent to the linear model. A full explanation of ANOVA is [here](https://mpfoley73.github.io/statistics/onewayanova.html), but with survey analysis, the equal variances assumption does not apply because of the weighting.

::: panel-tabset

### Plot

```{r}
apistrat |>
  ggplot(aes(x = hsg_cut, y = api00)) +
  geom_boxplot()
```

### Model

```{r}
svyglm(
  api00 ~ hsg_cut,
  design = apistrat_des
) |>
  gtsummary::tbl_regression(intercept = TRUE)
```

:::

## Linear Regression

::: panel-tabset

### Plot

```{r}
apistrat |>
  ggplot(aes(x = api99, y = api00)) +
  geom_point() +
  facet_wrap(vars(hsg_cut))
```

### Model

```{r}
svyglm(
  api00 ~ api99 + hsg_cut,
  design = apistrat_des
) |>
  gtsummary::tbl_regression(intercept = TRUE)
```

:::

## Logistic Regression

::: panel-tabset

### Plot

```{r}
apistrat_des |>
  summarize(
    .by = meals_cut,
    met_pct = survey_mean(
      sch.wide == "Yes", na.rm = TRUE, proportion = TRUE, vartype = "ci"
    )
  ) |>
  ggplot(aes(x = meals_cut, y = met_pct)) +
  geom_col(width = .4) +
  geom_errorbar(aes(ymin = met_pct_low, ymax = met_pct_upp), width = .1)
```

### Model

```{r}
#| label: tbl-model-logistic

svyglm(
  sch.wide ~ api99 + meals_cut,
  design = apistrat_des,
  family = quasibinomial
) |>
  gtsummary::tbl_regression(intercept = TRUE) |>
  as_gt() |>
  gt::tab_caption("Logistic regression of meeting school-wide target by API in 1999.")
```

:::
